Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-17 article-models-tt.tex
adding ideas to the introduction of section 4

\section{Type theory before type constructors}\label{tt}
In this section we give a description of type theory before basic
constructors and with explicit extension, weakening and substitution. We will
formulate the type theory in such a way that contexts aren't lists of variable
declarations. The main reason is that we don't see variable declarations in the
models either. This way we also set out to a more algebraic approach of type
theory and higher category theory. Thirdly, we will not have to be burdened with
comments about variables being bounded or not, or fresh or not occuring at all.

We will develop a theory which is in fact a theory of contexts and families of
contexts where a family of context can be a type, behind which we have the
intuition that it is an atomic or irreducible family of contexts. We get immediately
that every construction can be made in an arbitrary context.

It should be noted that this is just a prototype description. It may be the case
that various other elements have to be added yet, such as terms
$\jterm{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\ctxext{A}{P}}}{i}$
including the fiber $\subst{x}{P}$ into the extended family $\ctxext{A}{P}$ over
$\Gamma$. Moreover, it hasn't been thoroughly tested whether the theory presented
here lives up to its promise.

>>>>

\section{Type theory before type constructors}\label{tt}
In this section we give a description of dependent type theory before type
constructors with explicit operations of extension, weakening and substitution. We will
formulate the type theory in such a way that contexts aren't defined
to be lists of variable
declarations. We rather take a syntactic approach in our treatment of contexts,
axiomatizing only what you can do with them and allowing for the freedom to
interpret the axiom in any desired way. 
The main reason is that we don't see variable declarations in the internal
models either. This way we also set out to a more algebraic approach of type
theory and higher category theory. Thirdly, we will not have to be burdened with
superficial comments about variables being bounded or not, or fresh or free or 
not occuring at all.

The theory we will develop here is a theory of contexts, families of contexts
and terms thereof. Families of contexts could have been named dependent contexts
or telescopes.

In our treatment, a context can be the empty context or it can be a binary
planar tree of which `the leaves are (families of) contexts'. 
The judgmental equality relation on contexts is an equivalence relation which 
expresses that binary planar
trees of contexts are judgmentally equal if their leaves are, taking only the
order of the leaves into account \emph{and not the actual shape of the three}.
The intuition behind this equivalence relation is indeed that unstructured
(i.e.~unbracketed) lists such as the lists of variable declarations which
usually appear in type theoretical syntax, may be regarded as contexts.

Besides contexts, families and terms there will also be a notion of `type in
a context'. Only a family of contexts over a context $\Gamma$ is eligible to
be a type in the context $\Gamma$ and we have the intuition that `being a type'
expresses the property of `being an atomic or irreducible family of contexts'.

We will formulate fairly strict rules governing the judgmental equalities,
expressing that extension, weakening and substitution are combatible with
each other in a judgmental manner. This does not, however, diminish the role
of isomorphisms or of homotopies could play in the theory once identity types
are added. Indeed, types could still have non-trivial identity relations and
the category of types in a certain context could genuinely display higher
categorical structure, or so we conjecture.
